---
title: "Chapter 9: Visualization with Layers"
subtitle: "Part 1"
author: "Introduction to Data Science"
format:
  revealjs:
    theme: default 
    slide-number: true
    chalkboard: 
      theme: whiteboard
      buttons: false
    preview-links: auto
    execute:
      warning: false
      echo: true
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| label: setup
#| include: false
#| #| eval: true
library(tidyverse)
```

## Review of Data Import

::: incremental
-   What function would you use to read a file where fields were
    separated with "\|"?

-   Apart from file, skip, and comment, what other arguments do
    `read_csv()` and `read_tsv()` have in common?

-   What are the most important arguments to `read_fwf()`?
:::

::: notes
Answers:

1.  `read_delim()`
2.  All of them are the same. They're virtually identical except for the
    default delimiter.
3.  Position and width (can specify either way). Show examples.
:::

## Example of Reading from a Fixed Width File {.smaller}

```{r}
#| eval: false
# use built-in text file called fwf-sample.txt
fwf_sample <- readr_example("fwf-sample.txt")
writeLines(read_lines(fwf_sample))
# A vector of field widths
read_fwf(fwf_sample, fwf_widths(c(20, 10, 12), c("name", "state", "ssn")))
# Paired vectors of start and end positions
read_fwf(fwf_sample, fwf_positions(c(1, 30), c(20, 42), c("name", "ssn")))
```

## Find the Problem

Q: What is wrong with each of the following inline CSV files?

```{r}
#| eval: false
read_csv("a,b\n1,2,3\n4,5,6")
read_csv("a,b,c\n1,2\n1,2,3,4")
read_csv("a,b\n\"1")
read_csv("a,b\n1,2\na,b")
read_csv("a,b\n1,t\nt,f")
read_csv("a;b\n1;3")
```

::: notes
Answers:

1.  two variables names but three columns of data, concatenates last two
    into one column ignoring comma between
2.  three variable names but 2 columns of data in first row and 4 in
    second; creates df with NA in first row and concatenates last two
    numbers in second row
3.  Only 1 observation but two names. The  character quotes the quote
    symbol so we have no beginning to the data. No data is seen. The
    first row is incomplete which seems to be problematic
4.  Works fine but both columns are coerced to character variables.
5.  Second column is recognized as logical because of t and f.
6.  One character variable with the non-syntactic name `a;b`
:::

## Outline

-   Grammar of graphics review
-   Aesthetic mappings and settings
-   Faceting

## New
