---
title: "Midterm Exam Review Topics"
subtitle: "EXam Date: February 9, 2024"
toc: true
format:
  html:
    self-contained: true
editor: visual
execute:
  warning: false
  eval: false
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(ggplot2)
library(ggthemes)
```

# Data Visualization: Chapters 1 and 9

Every **ggplot2** plot has three key components:

1.  data,
2.  A set of aesthetic mappings between variables in the data and visual properties, and
3.  At least one layer which describes how to render each observation. Layers are usually created with a geom function.

Here’s a simple example:

```{r eval=FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point()
```

1.  Be familiar with the following geoms and their typical aesthetic specifications:

-   `geom_bar()`
-   `geom_boxplot()`
-   `geom_col()`
-   `geom_histogram()`
-   `geom_jitter()`
-   `geom_line()`
-   `geom_point()`
-   `geom_smooth()`

2.  Know common aesthetics and their relationship to type of variable (categorical vs quantitative/continuous) :

-   `x` and `y`
-   `color` and `fill`
-   `size`, `shape`, `alpha` and `linetype`

3.  Stats and their connections to geoms, especially for categorical variable plots like `geom_bar()`.

4.  Position adjustments and their relationship to `geom`s, especially for `geom_bar()` and `geom_point()`

5.  Faceting methods and their purposes:

    -   `facet_wrap()`and `facet_grid()`
    -   Use of nesting, e.g., \`facet_wrap(\~xvar + yvar)

6.  Use of `coord_flip()` and `coord_cartesian()` to control the appearance of plots.

# Data Transformation: Chapter 3

1.  Be familiar with the different types of variables in R.

2.  Know the purpose and syntax of the key **dplyr** functions:

-   `filter()`
-   `arrange()`
-   `select()`
-   `mutate()`
-   `group_by()`
-   `summarize()`

3.  Be able to describe the role of `group_by()` to change the scope of each function from operating on the entire dataset to operating on it group-by-group.

4.  Be familiar with "helper" functions for `select()` and `summarize()`. These need not be memorized since you have them on the cheat sheets.

5.  Other commands from this section you should know about and be able to use when necessary:

-   `distinct()`
-   `count()`
-   `slice_()` functions
-   `rename()`
-   `relocate()`\

# Data Tidying: Chapter 5

1.  Know what it means for a dataset to be tidy and be able to distinguish between a tidy and untidy dataset.
2.  Know the primary arguments for `pivot_longer()` and `pivot_wider()`
    -   `pivot_longer()`: `cols`, `names_to` and `values_to`
    -   `pivot_wider()`: `id_cols`, `names_from`, `values_from`
3.  

# Data Import: Chapter 7

# Study suggestions

-   Review reading

-   Review notes

-   Review group work and homework assignments

-   Work end of section problems in [R4DS2e](https://r4ds.hadley.nz/) and check partial list of solutions [here](https://mine-cetinkaya-rundel.github.io/r4ds-solutions/).

# Types of Questions

1.  Multiple choice: Which code chunk yields this output?
2.  What does this code do? What object(s) will it produce (e.g., dataset with 20 rows and 3 columns)
3.  Find the error and fix it (nothing very obscure).
4.  What goes in the missing spot in the code chunk?
5.  Complete the code chunk to create the given output.
