---
title: "Midterm Exam Review Topics"
subtitle: "EXam Date: February 9, 2024"
toc: true
format:
  html:
    self-contained: true
editor: visual
execute:
  warning: false
  eval: false
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(ggplot2)
library(ggthemes)
```



# Data Visualization: Chapters 1 and 9

Every **ggplot2** plot has three key components:

1. data,
2. A set of aesthetic mappings between variables in the data and visual properties, and
3. At least one layer which describes how to render each observation. Layers are usually created with a geom function.

Here’s a simple example:

```{r eval=FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point()
```

1. Be familiar with the following geoms and their typical aesthetic specifications:  

  * `geom_bar()`  
  * `geom_boxplot()`  
  * `geom_col()`  
  * `geom_density()`  
  * `geom_freqpoly()`  
  * `geom_histogram()`  
  * `geom_jitter()`  
  * `geom_line()`
  * `geom_point()`  
  * `geom_smooth()`  

2. Common aesthetics and their relationship to type of variable (categorical vs quantitative/continuous) :  

  * `x` and `y`  
  * `color` and `fill`  
  * `size`, `shape`, and `linetype`  

  
3. Scales and how they can be used and specified manually  

4. Stats and their connections to geoms, especially for categorical variable plots like `geom_bar()`.  

5. Position adjustments and their relationship to geoms  

6. Faceting methods and their purposes  

7. Use of `coord_flip()` and `coord_cartesian()` to control the appearance of plots.  

8. Practice on the end of section questions in R4DS.  I have a link to R4DS solutions near the top of the Moodle page but [it's also here](https://jrnold.github.io/r4ds-exercise-solutions/). These solutions also have some very valuable insights and extensions to the text material.  



# Data Transformation: Chapter 3

1. Be familiar with the different types of variables in R.  

2. Know the purpose and syntax of the five key **dplyr** functions:  
  * `filter()`  
  * `arrange()`  
  * `select()`  
  * `mutate()`  
  * `summarise()`  
  
3. Be able to describe the role of `group_by()` to change the scope of each function from operating on the entire dataset to operating on it group-by-group.  

4. Know how missing values work and how to use the `is.na()` function to include or remove cases with NA values.  

5. Be familiar with "helper" functions for `select()` and `summarize()`. These need not be memorized since you have them on the cheat sheets.  

6. Other commands from this section you should know about and be able to use when necessary:  

  * `distinct()`  
  * `count()`  
  * `slice()`  
  * `pull()`  
  * `top_n()`  
  
  
# Data Tidying: Chapter 5

# Data Import: Chapter 7

  
# Types of Questions  

1. Multiple choice: Which code chunk yields this output?  
2. What does this code do? What object(s) will it produce (e.g., dataset with 20 rows and 3 columns)
3. Find the error and fix it (nothing very obscure).  
4. What goes in the missing spot in the code chunk?  
5. Complete the code chunk to create the given output.  

